---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidytitanic

<!-- badges: start -->
<!-- badges: end -->

You can install development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("EvaMaeRey/tidytitanic")
```


```{r example}
library(tidytitanic)
tidy_titanic |> head()

flat_titanic |> head()

passengers |> head()
```


```{r}
library(tidytitanic)

head(flat_titanic)
head(tidy_titanic)
head(passengers)
```



### What's inside, A. Prepping data

For this package, preparation happens in the `data-raw` folder, in a file called `dataset_prep`.  

```{r}
flat_titanic <- Titanic |>
  data.frame() |>
  janitor::clean_names() |>
  tibble::tibble()

usethis::use_data(flat_titanic, overwrite = T)

library(tidyverse)
tidy_titanic <- Titanic |>
  data.frame() |>
  janitor::clean_names()|>
  uncount(weights = freq) |>
  mutate(id = 1:n(), .before = 1) |>
  remove_rownames() |>
  tibble()

usethis::use_data(tidy_titanic, overwrite = T)


passengers <-  
  dplyr::bind_rows(titanic::titanic_train, 
                   titanic::titanic_test |> 
                     dplyr::left_join(titanic::titanic_gender_model)
          ) |>
  janitor::clean_names() 

usethis::use_data(passengers, overwrite = T)
```


# document data

```{r data_documentation}


#' Tallied characteristics
#'
#' A dataset tallying frequencies of titanic passenger characteristics based on Titanic data from base R.  Includes Crew
#'
#' @format A data frame with 32 rows and 5 variables:
#' \describe{
#'   \item{class}{class (1st, 2nd, 3rd or Crew)}
#'   \item{sex}{sex of passenger}
#'   \item{age}{child or adult}
#'   \item{survived}{survival outcome of passanger}
#'   \item{freq}{tally of all in category}
#'   ...
#' }
#' @source R stats
"flat_titanic"


#' 
#'
#' A dataset titanic passengers and characteristics based on Titanic data from base R.  Includes Crew
#'
#' @format A data frame with 2201 rows and 5 variables:
#' \describe{
#'   \item{id}{an id for each passenger}
#'   \item{class}{class (1st, 2nd, 3rd or Crew)}
#'   \item{sex}{sex of passenger}
#'   \item{age}{child or adult}
#'   \item{survived}{survival outcome of passanger}

#'   ...
#' }
#' @source R stats
"tidy_titanic"



#' Titanic test data.
#'
#' @format Data frame with columns
#' \describe{
#' \item{passenger_id}{Passenger ID}
#' \item{survived}{survival status}
#' \item{pclass}{Passenger Class}
#' \item{name}{Name}
#' \item{sex}{Sex}
#' \item{age}{Age}
#' \item{sib_sp}{Number of Siblings/Spouses Aboard}
#' \item{parch}{Number of Parents/Children Aboard}
#' \item{ticket}{Ticket Number}
#' \item{fare}{Passenger Fare}
#' \item{cabin}{Cabin}
#' \item{embarked}{Port of Embarkation}
#' 
#' ...
#' }
#' @source titanic package, https://www.kaggle.com/c/titanic/data
"passengers"
```

titanic is an R package containing data sets providing information on the fate of passengers on the fatal maiden voyage of the ocean liner "Titanic", with variables such as economic status (class), sex, age and survival. These data sets are often used as an introduction to machine learning on Kaggle. More details about the competition can be found here, and the original data sets can be found here.

```{r}
knitrExtra::chunk_to_dir("data_documentation", dir = "R")
```


```{r, eval = F}
devtools::document()
devtools::check()
devtools::install(".", upgrade = "never")
devtools::submit_cran()
```



